---
title: "Lab4"
author: "Tiana Chen"
date: "2024-09-26"
output: "github_document"
---

```{r}
library(car)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)


```

# Load dataset
```{r}

lab4data <- read.csv("/Users/chenxuejing/Documents/GitHub/Lab4/skewed.csv")


```


# Normality
## Normality Plots
```{r}


ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2)  + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()


#functions below are base R functions, they can do the job, but not as customizable as ggplot

hist(lab4data$Performance)

qqnorm(lab4data$Performance, col = "steelblue", lwd = 2)

plot(density(lab4data$Performance, na.rm = TRUE, bw = 90),  lwd=2, main = "")

```

### Normality Plots by Group
```{r}

#a fun function to plot the violin plot by group, but it's not part of ggplot
violinBy(Performance ~ Group, data = lab4data, rain= TRUE, vertical = FALSE)


#Find a way to plot the histograms, density, and qq plots by groups using ggplot
ggplot(lab4data, aes(x = Performance)) + geom_histogram(binwidth = 20) +  facet_wrap(~Group) + theme_classic()

ggplot(lab4data, aes(x = Performance)) + geom_density(adjust = 2) +  facet_wrap(~Group) + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()

#Use ggplot to create a violin plot by groups


```


## Normality Tests
```{r}

describe(lab4data$Performance)

shapiro.test(lab4data$Performance)



```

### Normality Tests by Group
```{r}

#Use the describeBy() function to get skewness and kurtosis by group
?describeBy()
describeBy(Performance ~ Group, data = lab4data)

#Use the group by function to get shapiro test results by group
lab4data %>%
  group_by(Group) %>%
  summarize(W = shapiro.test(Performance)$statistic,  p_value = shapiro.test(Performance)$p.value)

#Use the filter function to get both





```
# Q1: What is your overall conclusion about this variable's normality? Why?
All of these groups are insignificant because their p-values are all larger than 0.05.


# Equal Variance between Groups
## Descrptive Variance
```{r}
lab4data_clean<-drop_na(lab4data)
var(lab4data_clean$Performance)

lab4data_clean%>%
  group_by(Group) %>%
  summarize(variance = var(Performance))


```

## Equal Variance Test
```{r}


leveneTest(Performance~Group, lab4data)

MANOVA(lab4data, dv = "Performance", between = "Group")

#What if you want to test equal variance between 2 groups specifically? 

lab4dataConG1<-lab4data %>%
  filter(Group == "Control" | Group == "G1")

leveneTest(Performance~Group, lab4dataConG1)


```

# Q2: Overall, does it meet the equal variance assumption across the groups? Why?
No, it doesn't meet the equal variance assumption across those specific groups. Because the variances of G1, G2, G3, and G4 are mroe than 2 times bigger than the variance of the control group.
# Transformation
```{r}

#if any of the assumption is not met, use transformation 

lab4data$Performance_log <- log10(lab4data$Performance)

ggplot(lab4data, aes(x = Performance_log)) + geom_histogram(binwidth = 0.25) +facet_wrap(~Group) + theme_classic()

ggplot(lab4data, aes(x = Performance_log)) + geom_density(adjust = 2) +  facet_wrap(~Group) + theme_classic()

qq<-ggplot(lab4data, aes(sample = Performance_log)) + geom_qq()  + theme_classic()

qq+ geom_qq_line()
```

# Q3: Run the above tests again with the transformed outcome. Compare the differences in results. 
Before Logging (first image in the earlier set): The data shows a right-skewed distribution, with most of the counts concentrated in the lower performance range. There is a long tail toward the higher performance values, indicating that there are a few very high values that are not well distributed.
After Logging (new histogram): The log transformation compresses the range of values, resulting in a distribution that appears more symmetrical and less skewed. The peak is more centered, and the higher performance values that were previously creating the long tail are now better distributed.


